---
title: "Code"
output: html_document
---

```{r}
install.packages("ggplot2")
install.packages("tidyr")
install.packages("janitor")
install.packages("dplyr")
install.packages("shiny")
install.packages("tidyverse")
install.packages("matrixStats")
install.packages("e1071")

rowSums() # sum of row
colSums() # sum of column
rowMeans()
colMeans()
rowSds()
colSds()
skewness()
kurtosis()

```

```{r}
library(ggplot2)
library(tidyr)
library(janitor)
library(dplyr)
library(shiny)
library(tidyverse)
library(matrixStats)
library(e1071)
```


```{r}
Students <- c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j", "k")

Course1 <- c(6.14, 8.20, 9.05, 5.54, 4.06, 3.14, 6.98, 8.55, 8.21, 9.94, 2.39)
Course2 <- c(7.70, 9.17, 9.77, 7.44, 9.41, 7.48, 6.13, 7.38, 7.53, 8.11, 9.41)
Course3 <- c(5.88, 8.28, 5.04, 8.49, 6.00, 8.62, 6.20, 8.13, 9.19, 9.43, 7.00)
Course4 <- c(2.80, 1.23, 7.98, 8.43, 1.81, 6.47, 3.40, 5.37, 3.39, 1.61, 0.72)
Course5 <- c(7.63, 5.19, 5.71, 6.60, 4.95, 6.91, 5.80, 6.83, 7.93, 5.35, 6.86)
Course6 <- c(8.43, 9.91, 8.58, 8.29, 10.00, 8.15, 8.93, 8.80, 8.61, 8.94, 9.59)
student_data <- data.frame(Students, Course1, Course2, Course3, Course4, Course5, Course6)
str(student_data)
view(student_data)


```
```{r}
student_data2 <- student_data[,-1] # converted the student names into row names
rownames(student_data2) <- student_data[,1]
view(student_data2)

# creating new columns to do calculations

SumperStudent <- rowSums(student_data2[,c("Course1","Course2","Course3","Course4","Course5","Course6")])

MeanperStudent <- rowMeans(student_data2[,c("Course1","Course2","Course3","Course4","Course5","Course6")])

SDperStudent <- apply(student_data2, 1, sd)

SumperStudent # sum score per student

SDperStudent # standard deviation per student

student_data2 %>% mutate(SumperStudent, MeanperStudent, SDperStudent) %>% head(5)

student_data2 # viewing the object in a separate window
```




```{r}
student_data$row_std <- rowSds(as.matrix(student_data[,c("Course1","Course2","Course3","Course4","Course5","Course6")]))

student_data$row_std <- apply(student_data[,-1], 1, sd) # focus on fixing this
student_data






```

```{r}
new_row <- data.frame(sum(Students), sum(Course1),sum(Course2),sum(Course3),sum(Course4),sum(Course5),sum(Course6), row.names="Sum")

new_row 

names(new_row) <- names(student_data)
student_data <- rbind(student_data, new_row)
tail(student_data)
```


```{r}
student_data <- student_data %>% mutate(Sumperstudent) %>% mutate(Meanperstudent)
```


```{r}

student_data$row_mean <- rowMeans(student_data[,c("Course1","Course2","Course3","Course4","Course5","Course6")])

student_data$row_mean <- apply(student_data[,-1], 1, mean)

student_data$row_sums <- rowSums(student_data[,c("Course1","Course2","Course3","Course4","Course5","Course6")])

student_data$row_sums <- apply(student_data[,-1], 1, sum)
```

creating functions and loops and statements
```{r}

MeanperStudentsTrue <- apply(student_data2[,1:6], 1, mean) # we applied the mean of all the courses of student a in a column
MeanperStudentsTrue


meanpercourse <- apply(student_data2[1:11,], 2, mean) # we applied the mean of all the courses of all the students in a vector and attached it to an object called
meanpercourse # it works


skewpercourse <- apply(student_data2[1:11,], 2, skewness) # we applied the skew of all the courses of student a in a column

skewpercourse # function needs to be created

sdpercourse <- apply(student_data2[1:11,1:6], 2, sd) # we applied the sd of all the courses of student a in a column
sdpercourse #  function needs to be created

# Kurtosis function needs to be created

add the bin, cumulative and frequency dataframe

```
```{r}
# this chunk is useless for now

for (number in 1:5){
  result[number] <- 25  # we can put anything here so we can use the SD, Kur and Skew calculations here
}


# an attempt to create loop function

 # result3 <- c()
 # 
 # for (number in 1:1){
 #   result[number] <- 6.14 - 6.43
 #   result <- result^2
 #   result2 <- mean(result2)
 #   result3 <- sqrt(result3)
 # }
 # 
 # result4
# sumCourse1 <- sum(Course1)
# sqdifcourse1 <- mean(Course1) 



```


Remove intersect values to NA?

```{r}
# fixed (but has to be updated once the dataframe is complete)

student_data2[12, 7] = NA # 1 single observation has been changed
view(student_data2)
student_data2$SumperStudents[12:16] = NA # certain part of a row
student_data2$MeanperStudents[12:16] = NA
student_data2$SdperStudents[12:16] = NA
student_data2$SkewperStudents[12:16] = NA
student_data2$KurtosisperStudents[12:16] = NA
view(student_data2)


```



Plotting!


## Create a histogram in numbers of marks
```{r}
ggplot(..., aes(x=.....)) + 
  geom_histogram() 
```

## Create a histogram in %
```{r}
ggplot(..., aes(x=.....)) + 
  geom_histogram() 
```

## Create a graph with the overall distribution of marks
```{r}
p1 <- ggplot(....., aes(x=....., y=......)) +
  geom_line(color="#69b3a2", size=2)
p2 <- ggplot(...., aes(x=...., y=.....)) +
  geom_line(color="grey",size=2)
p1 + p2 
```

## Create a graph with the overall distribution of marks in %
```{r}
p3 <- ggplot(....., aes(x=....., y=......)) +
  geom_line(color="#69b3a2", size=2)
p4 <- ggplot(...., aes(x=...., y=.....)) +
  geom_line(color="grey",size=2)
p3 + p4
```

## Create a graph with the mean grade by student
```{r}
ggplot(....., aes(x=....., y=......)) +
  geom_line()
```

## Create a graph with the mean grade by course
```{r}
ggplot(....., aes(x=....., y=......)) +
  geom_line()
```

## Create a graph with the std of grades by student
```{r}
ggplot(....., aes(x=....., y=......)) +
  geom_line()
```

## Create a graph with the std of grades by course
```{r}
ggplot(....., aes(x=....., y=......)) +
  geom_line()
```

## Create a graph with the skewness of grades by student
```{r}
ggplot(....., aes(x=....., y=......)) +
  geom_line()
```

## Create a graph with the skewness of grades by course
```{r}
ggplot(....., aes(x=....., y=......)) +
  geom_line()
```

## Create a graph with the kurtosis of grades by student
```{r}
ggplot(....., aes(x=....., y=......)) +
  geom_line()
```

## Create a graph with the kurtosis of grades by course
```{r}
ggplot(....., aes(x=....., y=......)) +
  geom_line()
```



